\section{Improvement Heuristics 2-OPT}

\subsection{The Algorithm}

Improvement heuristics are the polar opposite of construction heuristics. They need an initial solution to start from. The solution is then modified to get a better neighboring solution.

In particular, 2-OPT takes a tour and swaps two edges to reduce the tour length. It keeps doing this until no such edge pair can be found. Algorithm~\ref{algo:2-opt} describes this process.

\begin{algorithm}
    \caption{2--OPT}
    \label{algo:2-opt}

    \KwIn{instance \comment{An instance of TSP} }
    \KwIn{starting\_tour \comment{A tour of the instance}}
    \SetKwFunction{range}{range}
    \SetKwFunction{distance}{distance}
    \SetKwFunction{swap}{swap}

    \Begin{
        \Repeat{exit}{
            exit \(gets\) true\;
            \ForEach(\comment{For evry city in the tour}){i \(\in \llbracket 0, n-1\rrbracket\)}{
                \ForEach(\comment{For evry city non-adjacent to i in the tour}){j in \(\in \llbracket 0, n-1\rrbracket- \{i - 1, i, i + 1\}\)}{
                    \If(\comment{If the swap is benificial}){\(\distance{i, i + 1} + \distance{j, j + 1} < \distance{i, j} + \distance{j + 1, i + 1}\)}{
                        \swap{i, i+1, j, j+1}\comment{Swap the edges}\;
                        exit \(gets\) false \comment{Keep looping}\;
                    }
                    
                }
            }
        }
    }
\end{algorithm}

\subsection{Performance Analysis}

From a runtime perspective, 2-OPT is a somewhat decent heuristic. Its average case runtime is \(O(n^2)\), but its worst case runtime is exponential. This makes it generally too expensive to use on its own, but it is very useful when combined with other heuristics.

The solutions it gives by contrast are of very high quality. Very tight upper bounds on the 2-OPT tour's length can be proven under the assumption of the triangle inequality~\cite{opt2-ratio}, and even better bounds can be achieved. This is formalized by Proposition~\ref{prop:opt2-ratio}.

\begin{proposition}[Bounds on 2--OPT]
    \label{prop:opt2-ratio}
    If \(x\) is a \TSP{} instance, then \[\mu_x(2OPT(x))\le \sqrt{n\over 2} \mathrm{opt}(x)\]
\end{proposition}